Forecasting biodiversity in breeding birds using best practices

### 2017-11-20 manuscript/draft.Rmd
table formatting

---
title: "Forecasting biodiversity in breeding birds using best practices"
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \usepackage{booktabs}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

@@

@@

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity is increasing, there is little evidence
showing that these forecasts work well at decadal time scales. Most biodiversity
forecasts are not evaluated to determine how well they predict future diversity,
fail to account for uncertainty, and do not use time-series data that captures
the actual dynamics being studied. In this study, we used best practices and
32-year time series to evaluate six different modeling approaches to predicting
changes in North American breeding birds' species richness. Hindcasts for each
method were evaluated annually for a decade at `r numbers$N_predict_sites` sites
distributed throughout the continental United States. While each model could
explain most of the variance in richness, none of them outperformed baseline
models that predicted constant richness at each site. In particular, we found no
evidence that current methods (such as species distribution models) can
successfully turn spatial data into useful temporal predictions about
biodiversity at decadal time-scales. To facilitate the rapid improvement of
biodiversity forecasts, we emphasize the value of specific best practices at
multiple stages of the forecasting pipeline, from the choice of predictor
variables through the choice of approach for evaluating model performance over
time.

@@

We evaluated 6 types of forecasting models (Table 1) by dividing the 32 years
of data into 22 years of training data and 10 years of data for 
evaluating forecasts using hindcasting. We also made long term forecasts by using
the full data set for training and making forecasts through the year 2050. For
both time scales, we made forecasts using each model with and without
correcting for observer effects, as described below.

@@

**Richness data.** Bird species richness was obtained from the North American
Breeding Bird Survey (BBS) [@pardieck2017] using the Data Retriever Python
package [@morris2013] and rdataretriever R package [@rdataretriever]. The BBS
data was filtered to exclude all nocturnal, cepuscular, and aquatic species
[since these species are not well sampled by BBS methods; @hurlbert2005], as
well as unidentified species, and hybrids. All data from surveys that did not
meet BBS quality criteria were also excluded.

@@

table_data = read_csv("../figures/table.csv",col_types = "ccccc")

table_data %>%
  knitr::kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    caption = table_caption
  ) %>%
  add_header_above(c(" " = 2, "Predictors" = 3)) %>%
  group_rows("Single-site models", 1, 3) %>%
  group_rows("Environmental models", 4, 6) %>% kable_styling(position = "center")
```

@@

**Time series models.** We used Auto-ARIMA models [based on the `auto.arima`
function in @forecast1] to represent an array of different time-series modeling
approaches. These models can include an autoregressive component (as in the
"naive" model, but with the possibility of longer-term dependencies in the
underlying process), a moving average component (where the noise can have serial
autocorrelation) and an integration/differencing component (so that the analysis
could be performed on sequential differences of the raw data, accommodating more
complex patterns including trends). The `auto.arima` function chooses whether to
include each of these components (and how many terms to include for each one)
using AICc [@forecast1]. Since there is no seasonal component to the BBS
time-series, we did not include a season component in these models. Otherwise we
used the default settings for this function [@forecast1].

@@

**Species Distribution Model: stacked random forests.** Species distribution
models (SDMs) predict individual species' occurrence probabilities using
environmental variables. Species-level models are used to predict richness by
summing the predicted probability of occupancy across all species at a site.
This avoids known problems with the use of thresholds for determining whether or
not a species will be present at a site [@calabrese2014; @pellissier2013].
Following @calabrese2014, we calculated the uncertainty in our richness estimate
by treating richness as a sum over independent Bernoulli random variables:
$\sigma^2_{richness} = \sum_i{p_i (1 - p_i)}$, where $i$ indexes species. By
itself, this approach is known to underestimate the true community-level
uncertainty because it ignores the uncertainty in the species-level probabilites
[@calabrese2014]. To mitigate this problem, we used an ensemble of 500 estimates
for each of the species-level probabilities instead of just one, propagating the
uncertainty forward. We obtained these estimates using random forests, a common
approach in the species distribution modeling literature. Random forests are
constructed by fitting hundreds of independent regression trees to
randomly-perturbed versions of the data [@cutler2007; @caruana2008]. When
correcting for observer effects, each of the 500 trees in our species-level
random forests used a different Monte Carlo estimate of the observer effects as
a predictor variable.

@@

All analyses were conducted using R [@R]. Primary R packages used in the
analysis included dplyr [@dplyr], tidyr [@tidyr], gimms [@gimms], sp [@sp1;
@sp2], raster [@raster], prism [@prism], rdataretriever [@rdataretriever],
forecast [@forecast1; @forecast2], git2r [@git2r], ggplot [@ggplot2], mistnet
[@harris2015], viridis [@viridis], rstan [@rstan], yaml [@yaml], purrr [@purrr],
gbm [Ridgeway *et al.* -@gbm], randomForest [@randomForest]. Code to fully
reproduce this analysis is available on GitHub
(https://github.com/weecology/bbs-forecasting) and archived on Zenodo
(https://doi.org/10.5281/zenodo.839581).

@@

Partitioning each model's squared error into consistent errors across sites and
errors that fluctuated year-to-year shows that the majority of the error comes
from errors in the site-level mean (Figure 6). The "average" model, which was
based on the site-level mean, had the lowest error in this regard. In contrast,
the three environmental models showed the largest biases at the site level; this
makes sense, given that they could not explicitly distinguish among sites with
similar climate, NDVI, and elevation.

@@

Accounting for differences among observers generally improved measures of model
fit (Figure 7). Improvements primarily resulted from a small number of forecasts
where observer turnover caused a large shift in the reported richness values
(Appendix [[Figure Number]]). The naive baseline was less sensitive to these
shifts, because it largely ignored the richness values reported by observers
that had retired by the end of the training period (Figure 1). The average
model, which gave equal weight to observations from the whole training period,
showed a larger decline in performance when not accounting for observer effects
-- especially in terms of coverage. The performance of the mistnet JSDM was
notable here, because its prediction intervals retained good coverage even when
not correcting for observer differences, which we attribute to the JSDM's
ability to model this variation with its latent variables.

@@

When comparing different methods for forecasting our results demonstrate the
importance of considering uncertainty [Box 1; @clark2001; @dietz2016]. Previous
comparisons between stacked SDMs and macroecological models reported that the
methods yielded equivalent results for forecasting diversity [@algar2009;
@distler2015]. While our results support this equivalence for point estimates,
they also show that stacked SDMs dramatically underestimate the range of
possible outcomes; after ten years, more than a third of the observed richness
values fell outside the stacked SDMs' 95% prediction intervals. Consistent with
@harris2015 and @warton2015, we found that JSDMs' wider prediction intervals
enabled them to avoid this problem. Macroecological models appear to share this
advantage, while being considerably easier to implement.

The accuracy for forecasts is expected to decline as forecasts are made further
into the future, so it is important to evaluate forecasts at different lead
times [Box 1; @petchey2015]. For deviance (the error metric that evaluates both
models' point estimates and their uncertainty simultaneously), there is some
indication that the JSDM and the nonlinear richness model converge towards the
time-series and baseline models as the lead time of the forecast increases. This
potential convergence is biologically plausible because at short lead times,
environmental changes are relatively small and changes in the biota may lag
behind changes in the environment. However, as the lead time of the forecast
increases the amount of environmental change should increase and the systems
will have more time to respond, potentially resulting in shifts towards improved
relative performance of models incorporating environmental and ecological
information [@blonder2017]. This is potentially important for interpreting our
results because we only evaluated annual forecasts up to a decade into the
future. In contrast, many ecological forecasts are temporally aggregated to 5-30
year timesteps and have lead times of up to a century. These long-term forecasts
are difficult to evaluate given the small number of century scale datasets, but
are important for understanding changes in biodiversity at some of the lead
times relevant for conservation and management. Two studies have assessed models
of species richness at longer lead times [@algar2009; @distler2015], but the
results were not compared to baseline or time-series models (in part due to data
limitations) making them difficult to compare to our results. Given the
potential changes with lead time of the models and processes are most relevant
to forecasting, it is important that more long lead time evaluation studies are
conducted and that the inferences drawn from short-term and long-term
forecasting studies be compared. For true forecasts, short term forecasts will
allow for relatively rapid validation of models by researchers without needing
to wait decades for data to accumulate [@tredennick2016; @dietze2016]. However,
since drivers of species richness differ at different temporal scales
[@rosenzweig1995; @white2004; @white2007], short-term forecasts should be viewed
as informing, not replacing, long-term forecasts [@tredennick2016].

While it is possible that models that include species' relationships to their
environments or direct environmental constraints on richness will provide better
fits at longer lead times, it is also possible that they will continue to
produce forecasts that are worse than baselines assuming that the systems are
static.  This would be expected to occur if richness in these systems is not
changing over the relevant multi-decadal time scales, which would make simpler
models with no directional change more appropriate. Recent suggestions that
local scale richness in some systems is not changing directionally at
multi-decadal scales supports this possibility [@brown2001; @ernest2001;
@vellend2013; @dornelas2014]. A lack of change in richness may be expected even
in the presence of substantial changes in environmental conditions and species
composition at a site due to replacement of species from the regional pool
[@brown2001; @ernest2001]. On average, the Breeding Bird Survey sites used in
this study show little change in richness [site-level SD of `r numbers$resid_sd`
species, after controlling for differences among observers; see also
@lasorte2005]. The absence of rapid change in this dataset is beneficial for the
absolute accuracy of forecasts across different sites: when a past year's
richness is already known, it is easy to estimate future richness.  The
site-level stability of the BBS data also explains why SDMs and macroecological
models perform relatively well at predicting future richness, despite failing to
capture changes in richness over time. However, this stability also makes it
difficult to improve forecasts relative to simple baselines, since those
baselines are already close to representing what is actually occurring in the
system. These results suggest that single-site models should be actively
considered for forecasts of richness and other stable aspects of
biodiversity. Our results also suggest that future efforts to understand and
forecast biodiversity should incorporate species composition, since lower-level
processes are expected to be more dynamic [@ernest2001; @dornelas2014] and
contain more useful information [@harris2015].

Future biodiversity forecasting efforts also need to address the uncertainty
introduced by the error in forecasting the environmental conditions that are
used as predictor variables. In this, and other hindcasting studies, the
environmental conditions for the "future" are known because the data has already
been observed. However, in real forecasts the environmental conditions
themselves have to be predicted, and environmental forecasts will also have
uncertainty and bias. Ultimately, ecological forecasts that use environmental
data will therefore be more uncertain than our current hindcasting efforts. It
is important to correctly incorporate this uncertainty in the predictor
variables into forecasting models [@clark2001; @dietze2017]. Limitations in
forecasting future environmental conditions, particularly at small scales, will
present continued challenges for models incorporating environmental variables,
and this may result in a continued advantage for simple single-site approaches.

In addition to comparing and improving the process models used for forecasting
it is important to consider the observation models.  When working with any
ecological dataset, there are imperfections in the sampling process that have
the potential to influence results. With large scale surveys and citizen science
datasets, such as the Breeding Bird Survey, these issues are potentially
magnified by the large number of different observers and by major differences in
the habitats and species being surveyed [@sauer1994]. Accounting for differences
in observers reduced the average error in our point estimates and also improved
the coverage of the confidence intervals. In addition, controlling for observer
effects resulted in changes in which models performed best, most notably
reducing the relative performance of the naive model's point estimates. This
suggests that, prior to accounting for observer effects, the naive model
performed well because it was capable of accommodating rapid shifts in estimated
richness introduced by changes in the observer. These kinds of rapid changes
were difficult for the other single-site models to accommodate, and so the
average performance of the ARIMA and average models improved once this source of
observation error was addressed. This demonstrates that modeling observation
error can be important for properly estimating and reducing uncertainty in
forecasts and can also lead to changes in the best methods for forecasting [Box
1]. Another key aspect of an ideal observation model is imperfect detection. In
this study, we did not address differences in detection probability across
species and sites [@boulinier1998] since there is no clear way to address this
issue using NA Breeding Bird Survey data without making strong assumptions about
the data [i.e., assuming there is no biological variation in stops along a
route; @white2010], but this would be a valuable addition to future forecasting
models.

The science of forecasting biodiversity remains in its infancy and it is
important to consider weaknesses in current forecasting methods in that
context. In the beginning weather forecasts were also worse than simple
baselines, but these forecasts have continually improved throughout the history
of the field [@mcgill2012; @silver2012; @bauer2015]. One practice that lead to
improvements in weather forecasts was that large numbers of forecasts were made
publicly, allowing different approaches to be regularly assessed and refined
[@mcgill2012; @silver2012]. To facilitate this kind of improvement, it is
important for ecologists to start regularly making and evaluating real
ecological forecasts, even if they perform poorly, and to make these forecasts
openly available for assessment [@mcgill2012; @dietze2016]. These forecasts
should include both short-term predictions, which can be assessed quickly, and
mid- to long-term forecasts, which can help ecologists to assess long time-scale
processes and determine how far into the future we can successfully forecast
[@dietze2016; @tredennick2016]. Forecasts out to 2050 based on the models in
this paper have been openly archived [@zenodo], so that we and others can assess
how well they perform. We plan to evaluate these forecasts and report the
results as each new year of BBS data becomes available, and make iterative
improvements to the forecasting models in response to these assessments.

>>>>

---
title: "Forecasting biodiversity in breeding birds using best practices"
author:
- David J. Harris$^1$ (corresponding author)
- Shawn D. Taylor$^2$
- Ethan P. White$^1$
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \usepackage{booktabs}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: 
    keep_tex: true
  html_document: default
---

$^1$ Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, FL, United States

$^2$ School of Natural Resources and Environment, University of Florida Gainesville, FL, United States

@@

\newpage

@@

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity is increasing, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated to determine how well they predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We addressed these limitations by using best practices
to explore our ability to forecast the species richness of breeding birds in
North America. We used hindcasting to evaluate six different modeling approaches
for predicting richness. Hindcasts for each method were evaluated annually for a
decade at `r numbers$N_predict_sites` sites distributed throughout the
continental United States. While each model could explain most of the variance
in richness, none of them consistently outperformed a baseline model that
predicted constant richness at each site. In particular, we found no evidence
that current methods (such as species distribution models) can successfully turn
spatial data into useful temporal predictions about biodiversity at decadal
time-scales. The best practices implemented in this study directly influence the
forecasts, the relative performance of different modeling approaches, and the
conclusions about the current state of biodiversity forecasting . To facilitate
the rapid improvement of biodiversity forecasts, we emphasize the value of
specific best practices in making forecasts and evaluating forecasting methods.

@@

We evaluated 6 types of forecasting models (Table 1) by dividing the 32 years
of data into 22 years of training data and 10 years of data for evaluating
forecasts using hindcasting. Here we use definitions from meteorology, where a
hindcast is generally any prediction for an event that has already happened,
while forecasts are predictions for actual future events [@jolliffe2003]. We
also made long term forecasts by using the full data set for training and
making forecasts through the year 2050. For both time frames, we made forecasts
using each model with and without correcting for observer effects, as described
below.

@@

**Richness data.** Bird species richness was obtained from the North American
Breeding Bird Survey (BBS) [@pardieck2017] using the Data Retriever Python
package [@morris2013] and rdataretriever R package [@rdataretriever]. BBS
observations are three-minute point counts made at 50 fixed locations along a
40km route. Here we denote each route as a site and summarize richness as the
total species observed at all 50 locations in each surveyed year. Prior to
summarizing the data was filtered to exclude all nocturnal, cepuscular, and
aquatic species [since these species are not well sampled by BBS methods;
@hurlbert2005], as well as unidentified species, and hybrids. All data from
surveys that did not meet BBS quality criteria were also excluded.

@@

table_data = read_csv("../figures/table.csv",col_types = "ccccc", na = "na")

table_data %>%
  knitr::kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    caption = table_caption,
    align = c("l", "l", "c", "c", "c")
  ) %>%
  add_header_above(c(" " = 2, "Predictors" = 3)) %>%
  group_rows("Single-site models", 1, 3) %>%
  group_rows("Environmental models", 4, 6) %>% 
  kable_styling(position = "center")
```

@@

**Time series models.** We used Auto-ARIMA models [based on the `auto.arima`
function in the package `forecast`; @forecast1] to represent an array of
different time-series modeling approaches. These models can include an
autoregressive component (as in the "naive" model, but with the possibility of
longer-term dependencies in the underlying process), a moving average component
(where the noise can have serial autocorrelation) and an
integration/differencing component (so that the analysis could be performed on
sequential differences of the raw data, accommodating more complex patterns
including trends). The `auto.arima` function chooses whether to include each of
these components (and how many terms to include for each one) using AICc
[@forecast1]. Since there is no seasonal component to the BBS time-series, we
did not include a season component in these models. Otherwise we used the
default settings for this function [@forecast1].

@@

**Species Distribution Model: stacked random forests.** Species distribution
models (SDMs) predict individual species' occurrence probabilities using
environmental variables. Species-level models are used to predict richness by
summing the predicted probability of occupancy across all species at a site.
This avoids known problems with the use of thresholds for determining whether or
not a species will be present at a site [@calabrese2014; @pellissier2013].
Following @calabrese2014, we calculated the uncertainty in our richness estimate
by treating richness as a sum over independent Bernoulli random variables:
$\sigma^2_{richness} = \sum_i{p_i (1 - p_i)}$, where $i$ indexes species. By
itself, this approach is known to underestimate the true community-level
uncertainty because it ignores the uncertainty in the species-level probabilites
[@calabrese2014]. To mitigate this problem, we used an ensemble of 500 estimates
for each of the species-level probabilities instead of just one, propagating the
uncertainty forward. We obtained these estimates using random forests
[@randomForest], a common approach in the species distribution modeling
literature. Random forests are constructed by fitting hundreds of independent
regression trees to randomly-perturbed versions of the data [@cutler2007;
@caruana2008]. When correcting for observer effects, each of the 500 trees in
our species-level random forests used a different Monte Carlo estimate of the
observer effects as a predictor variable.

@@

All analyses were conducted using R [@R]. Primary R packages used in the
analysis included dplyr [@dplyr], tidyr [@tidyr], gimms [@gimms], sp [@sp1;
@sp2], raster [@raster], prism [@prism], rdataretriever [@rdataretriever],
forecast [@forecast1; @forecast2], git2r [@git2r], ggplot [@ggplot2], mistnet
[@harris2015], viridis [@viridis], rstan [@rstan], yaml [@yaml], purrr [@purrr],
gbm [Ridgeway *et al.* -@gbm], randomForest [@randomForest]. Code to fully
reproduce this analysis is available on GitHub
(https://github.com/weecology/bbs-forecasting) and archived on Zenodo
[@zenodoCode].

@@

Partitioning each model's squared error shows that the majority of the residual
error was attributed to errors in estimating site-level means, rather than
errors in tracking year-to-year fluctuations (Figure 6). The "average" model,
which was based entirely on site-level means, had the lowest error in this
regard. In contrast, the three environmental models showed larger biases at the
site level, though they still explained most of the variance in this component.
This makes sense, given that they could not explicitly distinguish among sites
with similar climate, NDVI, and elevation. Interestingly, the environmental
models had higher squared error than the baselines did for tracking year-to-year
fluctuations in richness as well.

@@

Accounting for differences among observers generally improved measures of model
fit (Figure 7). Improvements primarily resulted from a small number of forecasts
where observer turnover caused a large shift in the reported richness values.
The naive baseline was less sensitive to these shifts, because it largely
ignored the richness values reported by observers that had retired by the end of
the training period (Figure 1). The average model, which gave equal weight to
observations from the whole training period, showed a larger decline in
performance when not accounting for observer effects -- especially in terms of
coverage. The performance of the mistnet JSDM was notable here, because its
prediction intervals retained good coverage even when not correcting for
observer differences, which we attribute to the JSDM's ability to model this
variation with its latent variables.

@@

When comparing different methods for forecasting our results demonstrate the
importance of considering uncertainty [Box 1; @clark2001; @dietze2016]. Previous
comparisons between stacked SDMs and macroecological models reported that the
methods yielded equivalent results for forecasting diversity [@algar2009;
@distler2015]. While our results support this equivalence for point estimates,
they also show that stacked SDMs dramatically underestimate the range of
possible outcomes; after ten years, more than a third of the observed richness
values fell outside the stacked SDMs' 95% prediction intervals. Consistent with
@harris2015 and @warton2015, we found that JSDMs' wider prediction intervals
enabled them to avoid this problem. Macroecological models appear to share this
advantage, while being considerably easier to implement.

We have only evaluated annual forecasts up to a decade into the future, but
forecasts are often made with a lead time of 50 years or more. These long-term
forecasts are difficult to evaluate given the small number of century-scale
datasets, but are important for understanding changes in biodiversity at some of
the lead times relevant for conservation and management. Two studies have
assessed models of species richness at longer lead times [@algar2009;
@distler2015], but the results were not compared to baseline or time-series
models (in part due to data limitations) making them difficult to compare to our
results directly. Studies on shorter time scales, such as ours, provide one way
to evaluate our forecasting methods without having to wait several decades to
observe the effects of environmental change on biodiversity [@petchey2015;
@tredennick2016; @dietze2016], but cannot fully replace longer-term evaluations
[@tredennick2016]. In general, drivers of species richness can differ at
different temporal scales [@rosenzweig1995; @white2004; @white2007;
@blonder2017], so different methods may perform better for different lead times.
In particular, we might expect environmental and ecological information to
become more important at longer time scales, and thus for the performance of
simple baseline forecasts to degrade faster than forecasts from SDMs and other
similar models.  We did observe a small trend in this direction: deviance for
the auto-ARIMA models and for the average baseline grew faster than for two of
the environmental models (the JSDM and the macroecological model), although this
growth was not statistically significant for the average baseline.

While it is possible that models that include species' relationships to their
environments or direct environmental constraints on richness will provide better
fits at longer lead times, it is also possible that they will continue to
produce forecasts that are worse than baselines that assume the systems are
static.  This would be expected to occur if richness in these systems is not
changing over the relevant multi-decadal time scales, which would make simpler
models with no directional change more appropriate. Recent suggestions that
local scale richness in some systems is not changing directionally at
multi-decadal scales supports this possibility [@brown2001; @ernest2001;
@vellend2013; @dornelas2014]. A lack of change in richness may be expected even
in the presence of substantial changes in environmental conditions and species
composition at a site due to replacement of species from the regional pool
[@brown2001; @ernest2001]. On average, the Breeding Bird Survey sites used in
this study show little change in richness [site-level SD of `r numbers$resid_sd`
species, after controlling for differences among observers; see also
@lasorte2005]. The absence of rapid change in this dataset is beneficial for the
absolute accuracy of forecasts across different sites: when a past year's
richness is already known, it is easy to estimate future richness. Ward et al.
(2014) found similar patterns in time series of fisheries stocks, where
relatively stable time series were best predicted by simple models and more
complex models were only beneficial with dynamic time series. The site-level
stability of the BBS data also explains why SDMs and macroecological models
perform relatively well at predicting future richness, despite failing to
capture changes in richness over time. However, this stability also makes it
difficult to improve forecasts relative to simple baselines, since those
baselines are already close to representing what is actually occurring in the
system. These results suggest that single-site models should be actively
considered for forecasts of richness and other stable aspects of biodiversity.
Our results also suggest that future efforts to understand and forecast
biodiversity should incorporate species composition, since lower-level processes
are expected to be more dynamic [@ernest2001; @dornelas2014] and contain more
useful information [@harris2015].

Future biodiversity forecasting efforts also need to address the uncertainty
introduced by the error in forecasting the environmental conditions that are
used as predictor variables. In this, and other hindcasting studies, the
environmental conditions for the "future" are known because the data has already
been observed. However, in real forecasts the environmental conditions
themselves have to be predicted, and environmental forecasts will also have
uncertainty and bias. Ultimately, ecological forecasts that use environmental
data will therefore be more uncertain than our current hindcasting efforts, and
it is important to correctly incorporate this uncertainty into our models
[@clark2001; @dietze2017]. Limitations in forecasting future environmental
conditions---particularly at small scales---will present continued challenges
for models incorporating environmental variables, and this may result in a
continued advantage for simple single-site approaches.

In addition to comparing and improving the process models used for forecasting
it is important to consider the observation models.  When working with any
ecological dataset, there are imperfections in the sampling process that have
the potential to influence results. With large scale surveys and citizen science
datasets, such as the Breeding Bird Survey, these issues are potentially
magnified by the large number of different observers and by major differences in
the habitats and species being surveyed [@sauer1994]. Accounting for differences
in observers reduced the average error in our point estimates and also improved
the coverage of the confidence intervals. In addition, controlling for observer
effects resulted in changes in which models performed best, most notably
improving most models' point estimates relative to the naive baseline. This
demonstrates that modeling observation error can be important for properly
estimating and reducing uncertainty in forecasts and can also lead to changes in
the best methods for forecasting [Box 1]. This suggests that, prior to
accounting for observer effects, the naive model performed well largely because
it was capable of accommodating rapid shifts in estimated richness introduced by
changes in the observer. These kinds of rapid changes were difficult for the
other single-site models to accommodate. Another key aspect of an ideal
observation model is imperfect detection. In this study, we did not address
differences in detection probability across species and sites [@boulinier1998]
since there is no clear way to address this issue using North American Breeding
Bird Survey data without making strong assumptions about the data [i.e.,
assuming there is no biological variation in stops along a route; @white2010],
but this would be a valuable addition to future forecasting models.

The science of forecasting biodiversity remains in its infancy and it is
important to consider weaknesses in current forecasting methods in that context.
In the beginning, weather forecasts were also worse than simple baselines, but
these forecasts have continually improved throughout the history of the field
[@mcgill2012; @silver2012; @bauer2015]. One practice that lead to improvements
in weather forecasts was that large numbers of forecasts were made publicly,
allowing different approaches to be regularly assessed and refined [@mcgill2012;
@silver2012]. To facilitate this kind of improvement, it is important for
ecologists to start regularly making and evaluating real ecological forecasts,
even if they perform poorly, and to make these forecasts openly available for
assessment [@mcgill2012; @dietze2016]. These forecasts should include both
short-term predictions, which can be assessed quickly, and mid- to long-term
forecasts, which can help ecologists to assess long time-scale processes and
determine how far into the future we can successfully forecast [@dietze2016;
@tredennick2016]. We have openly archived forecasts from all six models through
the year 2050 [@zenodoPredict], so that we and others can assess how well they
perform. We plan to evaluate these forecasts and report the results as each new
year of BBS data becomes available, and make iterative improvements to the
forecasting models in response to these assessments.
